The socialization of meritocracy and market justice preferences at school

### 2024-10-02 submission/MDPI/paper-blinded.tex
Update revised version MPI Submit

\end{ideuno}
% -------------------
% END TITLE SECTION
% -------------------

@@

Table~\ref{tbl-lineal-reg} presents the results of linear mixed-effects
models for the market justice preference index. Model 1 shows that
higher perceived societal meritocracy, through effort, talent, and
deservingness, is positively and significantly associated with market
justice preferences. School meritocracy perceptions yield mixed effects:
perceived talent increases market justice preference, while perceived
effort decreases it, both statistically significant (\emph{p}
\textless{} 0.001). Model 2 includes family status, finding no
statistically significant effects from parents' education or the number
of books at home. Regarding school characteristics, Models 3 and 4
introduce socioeconomic level and academic performance. The results of
Model 3 suggest a statistically significant relationship between school
socioeconomic levels and market justice preferences, specifically those
schools with high socioeconomic levels (\(\beta\) = 0.37) compared to
low socioeconomic levels. However, Model 4 indicates that schools with
medium (\(\beta\) = -0.11) and high (\(\beta\) = -0.26) academic
performance present lower market justice preferences compared to
low-performing schools, with these effects being statistically
significant when controlling for other variables.

@@

The core of this paper was the association between meritocratic
perceptions and preferences for market justice. In general, we find that
those who perceive that effort and talent/intelligence are rewarded, are
more willing to agree with that richer individuals can have better
health, education, and pensions. This is not the first time that
meritocratic perceptions (and also beliefs) have been related with the
legitimation of social inequalities
{[}\citeproc{ref-mijs_paradox_2019}{49},\citeproc{ref-darnon_where_2018}{63}{]},
but the consideration of school level population as well as market
justice preferences allows adding evidence regarding the role of
meritocracy regarding inequality beliefs: legitimation of market
criteria in social services seems to begin at school age, and what
students think of meritocracy is linked to it. Although there are still
several open questions, as for instance differences between meritocratic
perceptions in school and in the society at large, this raises up the
relevance of what is done (or not) in schools in order to challenge the
meritocratic ideals.

@@

\bibitem[\citeproctext]{ref-chafel_schooling_1997}
\CSLLeftMargin{11. }%
\CSLRightInline{Chafel, J.A. Schooling, the {Hidden Curriculum}, and
{Children}'s {Conceptions} of {Poverty}. \emph{Social Policy Report}
\textbf{1997}, \emph{11}, 1--28,
doi:\href{https://doi.org/10.1002/j.2379-3988.1997.tb00004.x}{10.1002/j.2379-3988.1997.tb00004.x}.}

\bibitem[\citeproctext]{ref-flores_top_2020}
\CSLLeftMargin{12. }%
\CSLRightInline{Flores, I.; Sanhueza, C.; Atria, J.; Mayer, R. Top
{Incomes} in {Chile}: {A Historical Perspective} on {Income Inequality},
1964--2017. \emph{Review of Income and Wealth} \textbf{2020}, \emph{66},
850--874,
doi:\href{https://doi.org/10.1111/roiw.12441}{10.1111/roiw.12441}.}

\bibitem[\citeproctext]{ref-chancel_world_2022}
\CSLLeftMargin{13. }%
\CSLRightInline{Chancel, L.; Piketty, T.; Saez, E.; Zucman, G. World
Inequality Report 2022 2022.}

\bibitem[\citeproctext]{ref-ffrench-davis_reformas_2018}
\CSLLeftMargin{14. }%
\CSLRightInline{Ffrench-Davis, R. \emph{{Reformas econ{ó}micas en Chile
1973-2017}}; TAURUS, 2018; ISBN 978-956-9635-21-2.}

\bibitem[\citeproctext]{ref-boccardo_30_2020}
\CSLLeftMargin{15. }%
\CSLRightInline{Boccardo, G. \emph{30 a{ñ}os de Privatizaciones En
{Chile}: Lo Que La Pandemia Revel{ó}}; Nodo XXI.; Santiago, 2020;}

\bibitem[\citeproctext]{ref-observatoriosocial_estadisticas_2024}
\CSLLeftMargin{16. }%
\CSLRightInline{Social, O. \emph{Estad{í}sticas {Encuesta Casen} 2022,
Secci{ó}n {Salud}}; Observatorio social, 2024;}

\bibitem[\citeproctext]{ref-superintendenciadepensiones_estadisticas_2024}
\CSLLeftMargin{17. }%
\CSLRightInline{Superintendencia de Pensiones \emph{Estad{í}sticas de
{Afiliados} En {AFP} En {Sistema} de {Pensiones}. {Abril} 2024};
Superintendencia de Pensiones, 2024;}

\bibitem[\citeproctext]{ref-ministeriodeeducacion_resumen_2023}
\CSLLeftMargin{18. }%
\CSLRightInline{Ministerio de Educación \emph{Resumen de Matr{í}cula Por
Establecimiento Educacional}; Datos abiertos Ministerio de Educaci{ó}n,
2023;}

\bibitem[\citeproctext]{ref-boltanski_new_2005}
\CSLLeftMargin{19. }%
\CSLRightInline{Boltanski, L.; Chiapello, E. \emph{The New Spirit of
Capitalism}; Verso: London ; New York, 2005; ISBN 978-1-85984-554-7.}

\bibitem[\citeproctext]{ref-streeck_citizens_2012}
\CSLLeftMargin{20. }%
\CSLRightInline{Streeck, W. Citizens as {Customers}: {Considerations} on
the {New Politics} of {Consumption}. \emph{New Left Review}
\textbf{2012}, \emph{76}.}

\bibitem[\citeproctext]{ref-kluegel_beliefs_1987}
\CSLLeftMargin{21. }%
\CSLRightInline{Kluegel, J.R.; Smith, E.R. \emph{Beliefs about
{Inequality}: {Americans}' {Views} of {What} Is and {What Ought} to Be};
Routledge: New York, 1987; ISBN 978-0-202-30327-7.}

\bibitem[\citeproctext]{ref-wegener_dominant_1995}
\CSLLeftMargin{22. }%
\CSLRightInline{Wegener, B.; Liebig, S. Dominant {Ideologies} and the
{Variation} of {Distributive Justice Norms}: {A Comparison} of {East}
and {West Germany}, and the {United States}. In \emph{Social {Justice}
and {Political Change}: {Public Opinion} in {Capitalist} and
{Post-Communist States}}; Kluegel, J.R., Mason, D.S., Wegener, B., Eds.;
Walter de Gruyter: New York, 1995; pp. 239--259 ISBN 978-3-11-086894-4.}

\bibitem[\citeproctext]{ref-homan_being_2017}
\CSLLeftMargin{23. }%
\CSLRightInline{Homan, P.; Valentino, L.; Weed, E. Being and {Becoming
Poor}: {How Cultural Schemas Shape Beliefs About Poverty}. \emph{Social
Forces} \textbf{2017}, \emph{95}, 1023--1048,
doi:\href{https://doi.org/10.1093/sf/sox007}{10.1093/sf/sox007}.}

\bibitem[\citeproctext]{ref-castillo_legitimacy_2011}
\CSLLeftMargin{24. }%
\CSLRightInline{Castillo, J.C. Legitimacy of {Inequality} in a {Highly
Unequal Context}: {Evidence} from the {Chilean Case}. \emph{Social
Justice Research} \textbf{2011}, \emph{24}, 314--340,
doi:\href{https://doi.org/10.1007/s11211-011-0144-5}{10.1007/s11211-011-0144-5}.}

\bibitem[\citeproctext]{ref-evans_family_2010}
\CSLLeftMargin{25. }%
\CSLRightInline{Evans, M.; Kelley, J.; Sikora, J.; Treiman, D. Family
Scholarly Culture and Educational Success: {Books} and Schooling in 27
Nations. \emph{Research in Social Stratification and Mobility}
\textbf{2010}, \emph{28}, 171--197,
doi:\href{https://doi.org/10.1016/j.rssm.2010.01.002}{10.1016/j.rssm.2010.01.002}.}

\bibitem[\citeproctext]{ref-jasso_how_1999}
\CSLLeftMargin{26. }%
\CSLRightInline{Jasso, G. How {Much Injustice} Is {There} in the
{World}? {Two New Justice Indexes}. \emph{American Sociological Review}
\textbf{1999}, \emph{64}, 133--168,
doi:\href{https://doi.org/10.1177/000312249906400110}{10.1177/000312249906400110}.}

\bibitem[\citeproctext]{ref-shariff_income_2016}
\CSLLeftMargin{27. }%
\CSLRightInline{Shariff, A.F.; Wiwad, D.; Aknin, L.B. Income {Mobility
Breeds Tolerance} for {Income Inequality}: {Cross-National} and
{Experimental Evidence}. \emph{Perspectives on Psychological Science}
\textbf{2016}, \emph{11}, 373--380,
doi:\href{https://doi.org/10.1177/1745691616635596}{10.1177/1745691616635596}.}

\bibitem[\citeproctext]{ref-centeno_arc_2012}
\CSLLeftMargin{28. }%
\CSLRightInline{Centeno, M.A.; Cohen, J.N. The {Arc} of {Neoliberalism}.
\emph{Annual Review of Sociology} \textbf{2012}, \emph{38}, 317--340,
doi:\href{https://doi.org/10.1146/annurev-soc-081309-150235}{10.1146/annurev-soc-081309-150235}.}

\bibitem[\citeproctext]{ref-harvey_breve_2015}
\CSLLeftMargin{29. }%
\CSLRightInline{Harvey, D. \emph{{Breve historia del neoliberalismo}};
Ediciones Akal: Madrid (Espa{ñ}a), 2015; ISBN 978-84-460-2517-7.}

\bibitem[\citeproctext]{ref-jensen_worlds_2008}
\CSLLeftMargin{30. }%
\CSLRightInline{Jensen, C. Worlds of Welfare Services and Transfers.
\emph{Journal of European Social Policy} \textbf{2008}, \emph{18},
151--162,
doi:\href{https://doi.org/10.1177/0958928707087591}{10.1177/0958928707087591}.}

\bibitem[\citeproctext]{ref-stoy_worlds_2014}
\CSLLeftMargin{31. }%
\CSLRightInline{Stoy, V. Worlds of {Welfare Services}: {From Discovery}
to {Exploration}. \emph{Social Policy \& Administration} \textbf{2014},
\emph{48}, 343--360,
doi:\href{https://doi.org/10.1111/spol.12006}{10.1111/spol.12006}.}

\bibitem[\citeproctext]{ref-salamon_marketization_1993}
\CSLLeftMargin{32. }%
\CSLRightInline{Salamon, L.M. The {Marketization} of {Welfare}:
{Changing Nonprofit} and {For-Profit Roles} in the {American Welfare
State}. \emph{Social Service Review} \textbf{1993}, \emph{67}, 16--39,
doi:\href{https://doi.org/10.1086/603963}{10.1086/603963}.}

\bibitem[\citeproctext]{ref-sivesind_changing_2017}
\CSLLeftMargin{33. }%
\CSLRightInline{Sivesind, K.H.
\href{https://doi.org/10.1007/978-3-319-55381-8_2}{The {Changing Roles}
of {For-Profit} and {Nonprofit Welfare Provision} in {Norway}, {Sweden},
and {Denmark}}. In \emph{Promoting {Active Citizenship}}; Sivesind,
K.H., Saglie, J., Eds.; Springer International Publishing: Cham, 2017;
pp. 33--74 ISBN 978-3-319-55380-1 978-3-319-55381-8.}

\bibitem[\citeproctext]{ref-lindh_public_2015}
\CSLLeftMargin{34. }%
\CSLRightInline{Lindh, A. Public {Opinion} Against {Markets}?
{Attitudes} Towards {Market Distribution} of {Social Services} -- {A
Comparison} of 17 {Countries}. \emph{Social Policy \& Administration}
\textbf{2015}, \emph{49}, 887--910,
doi:\href{https://doi.org/10.1111/spol.12105}{10.1111/spol.12105}.}

\bibitem[\citeproctext]{ref-greitemeyer_subjective_2016}
\CSLLeftMargin{35. }%
\CSLRightInline{Greitemeyer, T.; Sagioglou, C. Subjective Socioeconomic
Status Causes Aggression: {A} Test of the Theory of Social Deprivation.
\emph{Journal of Personality and Social Psychology} \textbf{2016},
\emph{111}, 178--194,
doi:\href{https://doi.org/10.1037/pspi0000058}{10.1037/pspi0000058}.}

\bibitem[\citeproctext]{ref-mishra_subjective_2015}
\CSLLeftMargin{36. }%
\CSLRightInline{Mishra, S.; Carleton, R.N. Subjective Relative
Deprivation Is Associated with Poorer Physical and Mental Health.
\emph{Social Science \& Medicine} \textbf{2015}, \emph{147}, 144--149,
doi:\href{https://doi.org/10.1016/j.socscimed.2015.10.030}{10.1016/j.socscimed.2015.10.030}.}

\bibitem[\citeproctext]{ref-smith_relative_2012}
\CSLLeftMargin{37. }%
\CSLRightInline{Smith, H.J.; Pettigrew, T.F.; Pippin, G.M.;
Bialosiewicz, S. Relative {Deprivation}: {A Theoretical} and
{Meta-Analytic Review}. \emph{Personality and Social Psychology Review}
\textbf{2012}, \emph{16}, 203--232,
doi:\href{https://doi.org/10.1177/1088868311430825}{10.1177/1088868311430825}.}

\bibitem[\citeproctext]{ref-power_deprivationprotest_2018}
\CSLLeftMargin{38. }%
\CSLRightInline{Power, S.A. The {Deprivation-Protest Paradox}: {How} the
{Perception} of {Unfair Economic Inequality Leads} to {Civic Unrest}.
\emph{Current Anthropology} \textbf{2018}, \emph{59}, 765--789,
doi:\href{https://doi.org/10.1086/700679}{10.1086/700679}.}

\bibitem[\citeproctext]{ref-wilson_role_2003}
\CSLLeftMargin{39. }%
\CSLRightInline{Wilson, C. The {Role} of a {Merit Principle} in
{Distributive Justice}. \emph{The Journal of Ethics} \textbf{2003},
\emph{7}, 277--314,
doi:\href{https://doi.org/10.1023/A:1024667228488}{10.1023/A:1024667228488}.}

\bibitem[\citeproctext]{ref-bucca_merit_2016}
\CSLLeftMargin{40. }%
\CSLRightInline{Bucca, M. Merit and Blame in Unequal Societies:
{Explaining Latin Americans}' Beliefs about Wealth and Poverty.
\emph{Research in Social Stratification and Mobility} \textbf{2016},
\emph{44}, 98--112,
doi:\href{https://doi.org/10.1016/j.rssm.2016.02.005}{10.1016/j.rssm.2016.02.005}.}

\bibitem[\citeproctext]{ref-chong_mystery_2008}
\CSLLeftMargin{41. }%
\CSLRightInline{Chong, A.; Ñopo, H. The {Mystery} of {Discrimination} in
{Latin America}. \emph{Econom{í}a} \textbf{2008}, \emph{8}, 79--107,
doi:\href{https://doi.org/10.1353/eco.0.0005}{10.1353/eco.0.0005}.}

\bibitem[\citeproctext]{ref-torche_intergenerational_2014}
\CSLLeftMargin{42. }%
\CSLRightInline{Torche, F. Intergenerational {Mobility} and
{Inequality}: {The Latin American Case}. \emph{Annual Review of
Sociology} \textbf{2014}, \emph{40}, 619--642,
doi:\href{https://doi.org/10.1146/annurev-soc-071811-145521}{10.1146/annurev-soc-071811-145521}.}

\bibitem[\citeproctext]{ref-salgado_inequality_2023}
\CSLLeftMargin{43. }%
\CSLRightInline{Salgado, M.; Castillo, J. Inequality and
{Stratification} in {Latin America}. In \emph{The {Oxford Handbook} of
{Social Stratification}}; Gangl, M., Platt, L., Polavieja, J.G., van de
Werfhorst, H.G., Eds.; Oxford University Press: 2023, 2023; p. 0 ISBN
978-0-19-753948-4.}

\bibitem[\citeproctext]{ref-molyneux_neoliberal_2008}
\CSLLeftMargin{44. }%
\CSLRightInline{Molyneux, M. The {``{Neoliberal Turn}''} and the {New
Social Policy} in {Latin America}: {How Neoliberal}, {How New}?
\emph{Development and Change} \textbf{2008}, \emph{39}, 775--797,
doi:\href{https://doi.org/10.1111/j.1467-7660.2008.00505.x}{10.1111/j.1467-7660.2008.00505.x}.}

\bibitem[\citeproctext]{ref-centrodeestudiospublicos_estudio_2024}
\CSLLeftMargin{45. }%
\CSLRightInline{Centro de Estudios Públicos \emph{Estudio {Nacional} de
{Opini{ó}n P{ú}blica Base Consolidada}}; Centro de Estudios P{ú}blicos,
2024;}

\bibitem[\citeproctext]{ref-goldthorpe_myth_2003}
\CSLLeftMargin{46. }%
\CSLRightInline{Goldthorpe, J. The Myth of Education-Based Meritocracy.
\emph{New Economy} \textbf{2003}, \emph{10}, 234--239,
doi:\href{https://doi.org/10.1046/j.1468-0041.2003.00324.x}{10.1046/j.1468-0041.2003.00324.x}.}

\bibitem[\citeproctext]{ref-castillo_multidimensional_2023}
\CSLLeftMargin{47. }%
\CSLRightInline{Castillo, J.C.; Iturra, J.; Maldonado, L.; Atria, J.;
Meneses, F. A {Multidimensional Approach} for {Measuring Meritocratic
Beliefs}: {Advantages}, {Limitations} and {Alternatives} to the {ISSP
Social Inequality Survey}. \emph{International Journal of Sociology}
\textbf{2023}, \emph{53}, 448--472,
doi:\href{https://doi.org/10.1080/00207659.2023.2274712}{10.1080/00207659.2023.2274712}.}

>>>>

\end{ideuno}

% -------------------
% END TITLE SECTION
% -------------------

@@

Table~\ref{tbl-lineal-reg} ppresents the results of linear mixed-effects
models for the market justice preference index. Model 1 shows that
higher perceived societal meritocracy, through effort (\(\beta\) =
0.08), talent (\(\beta\) = 0.07), and deservingness (\(\beta\) = 0.2),
is positively and significantly associated with market justice
preferences (\emph{p} \textless{} 0.001). School meritocracy perceptions
yield mixed effects: perceived talent increases market justice
preference (\(\beta\) = 0.1), while perceived effort decreases it
(\(\beta\) = -0.12), both statistically significant (\emph{p}
\textless{} 0.001). Model 2 includes family status, finding no
statistically significant effects from parents' education or the number
of books at home. Regarding school characteristics, Models 3 and 4
introduce socioeconomic level and academic performance. The results of
Model 3 suggest a statistically significant relationship between school
socioeconomic levels and market justice preferences, specifically those
schools with high socioeconomic levels (\(\beta\) = 0.37; \emph{p}
\textless{} 0.01) compared to low socioeconomic levels. However, Model 4
indicates that schools with medium (\(\beta\) = -0.11; \emph{p}
\textless{} 0.01) and high (\(\beta\) = -0.26; \emph{p} \textless{}
0.001) academic performance present lower market justice preferences
compared to low-performing schools, with these effects being
statistically significant when controlling for other variables.

@@

The core of this paper was the association between meritocratic
perceptions and preferences for market justice. In general, we find that
those who perceive that effort and talent/intelligence are rewarded, are
more willing to agree with that richer individuals can have better
health, education, and pensions. This is not the first time that
meritocratic perceptions (and also beliefs) have been related with the
legitimation of social inequalities
{[}\citeproc{ref-mijs_paradox_2019}{51},\citeproc{ref-darnon_where_2018}{63}{]},
but the consideration of school level population as well as market
justice preferences allows adding evidence regarding the role of
meritocracy regarding inequality beliefs: legitimation of market
criteria in social services seems to begin at school age, and what
students think of meritocracy is linked to it. Although there are still
several open questions, as for instance differences between meritocratic
perceptions in school and in the society at large, this raises up the
relevance of what is done (or not) in schools in order to challenge the
meritocratic ideals.

@@

\bibitem[\citeproctext]{ref-lindh_public_2015}
\CSLLeftMargin{11. }%
\CSLRightInline{Lindh, A. Public {Opinion} Against {Markets}?
{Attitudes} Towards {Market Distribution} of {Social Services} -- {A
Comparison} of 17 {Countries}. \emph{Social Policy \& Administration}
\textbf{2015}, \emph{49}, 887--910,
doi:\href{https://doi.org/10.1111/spol.12105}{10.1111/spol.12105}.}

\bibitem[\citeproctext]{ref-castillo_meritocracia_2019}
\CSLLeftMargin{12. }%
\CSLRightInline{Castillo, J.C.; Torres, A.; Atria, J.; Maldonado, L.
Meritocracia y Desigualdad Econ{ó}mica: {Percepciones}, Preferencias e
Implicancias. \emph{Revista Internacional de Sociolog{í}a}
\textbf{2019}, \emph{77}, 117,
doi:\href{https://doi.org/10.3989/ris.2019.77.1.17.114}{10.3989/ris.2019.77.1.17.114}.}

\bibitem[\citeproctext]{ref-reynolds_perceptions_2014}
\CSLLeftMargin{13. }%
\CSLRightInline{Reynolds, J.; Xian, H. Perceptions of Meritocracy in the
Land of Opportunity. \emph{Research in Social Stratification and
Mobility} \textbf{2014}, \emph{36}, 121--137,
doi:\href{https://doi.org/10.1016/j.rssm.2014.03.001}{10.1016/j.rssm.2014.03.001}.}

\bibitem[\citeproctext]{ref-garcia-sierra_dark_2023}
\CSLLeftMargin{14. }%
\CSLRightInline{García-Sierra, A. The Dark Side of Meritocratic Beliefs:
{Is} Believing in Meritocracy Detrimental to Individuals from Low
Socioeconomic Backgrounds? \emph{Social Justice Research} \textbf{2023},
\emph{36}, 385--409,
doi:\href{https://doi.org/10.1007/s11211-023-00413-x}{10.1007/s11211-023-00413-x}.}

\bibitem[\citeproctext]{ref-flores_top_2020}
\CSLLeftMargin{15. }%
\CSLRightInline{Flores, I.; Sanhueza, C.; Atria, J.; Mayer, R. Top
{Incomes} in {Chile}: {A Historical Perspective} on {Income Inequality},
1964--2017. \emph{Review of Income and Wealth} \textbf{2020}, \emph{66},
850--874,
doi:\href{https://doi.org/10.1111/roiw.12441}{10.1111/roiw.12441}.}

\bibitem[\citeproctext]{ref-chancel_world_2022}
\CSLLeftMargin{16. }%
\CSLRightInline{Chancel, L.; Piketty, T.; Saez, E.; Zucman, G. World
Inequality Report 2022 2022.}

\bibitem[\citeproctext]{ref-ffrench-davis_reformas_2018}
\CSLLeftMargin{17. }%
\CSLRightInline{Ffrench-Davis, R. \emph{{Reformas econ{ó}micas en Chile
1973-2017}}; TAURUS, 2018; ISBN 978-956-9635-21-2.}

\bibitem[\citeproctext]{ref-boccardo_30_2020}
\CSLLeftMargin{18. }%
\CSLRightInline{Boccardo, G. \emph{30 a{ñ}os de Privatizaciones En
{Chile}: Lo Que La Pandemia Revel{ó}}; Nodo XXI.; Santiago, 2020;}

\bibitem[\citeproctext]{ref-observatoriosocial_estadisticas_2024}
\CSLLeftMargin{19. }%
\CSLRightInline{Social, O. \emph{Estad{í}sticas {Encuesta Casen} 2022,
Secci{ó}n {Salud}}; Observatorio social, 2024;}

\bibitem[\citeproctext]{ref-superintendenciadepensiones_estadisticas_2024}
\CSLLeftMargin{20. }%
\CSLRightInline{Superintendencia de Pensiones \emph{Estad{í}sticas de
{Afiliados} En {AFP} En {Sistema} de {Pensiones}. {Abril} 2024};
Superintendencia de Pensiones, 2024;}

\bibitem[\citeproctext]{ref-ministeriodeeducacion_resumen_2023}
\CSLLeftMargin{21. }%
\CSLRightInline{Ministerio de Educación \emph{Resumen de Matr{í}cula Por
Establecimiento Educacional}; Datos abiertos Ministerio de Educaci{ó}n,
2023;}

\bibitem[\citeproctext]{ref-boltanski_new_2005}
\CSLLeftMargin{22. }%
\CSLRightInline{Boltanski, L.; Chiapello, E. \emph{The New Spirit of
Capitalism}; Verso: London ; New York, 2005; ISBN 978-1-85984-554-7.}

\bibitem[\citeproctext]{ref-streeck_citizens_2012}
\CSLLeftMargin{23. }%
\CSLRightInline{Streeck, W. Citizens as {Customers}: {Considerations} on
the {New Politics} of {Consumption}. \emph{New Left Review}
\textbf{2012}, \emph{76}.}

\bibitem[\citeproctext]{ref-kluegel_beliefs_1987}
\CSLLeftMargin{24. }%
\CSLRightInline{Kluegel, J.R.; Smith, E.R. \emph{Beliefs about
{Inequality}: {Americans}' {Views} of {What} Is and {What Ought} to Be};
Routledge: New York, 1987; ISBN 978-0-202-30327-7.}

\bibitem[\citeproctext]{ref-wegener_dominant_1995}
\CSLLeftMargin{25. }%
\CSLRightInline{Wegener, B.; Liebig, S. Dominant {Ideologies} and the
{Variation} of {Distributive Justice Norms}: {A Comparison} of {East}
and {West Germany}, and the {United States}. In \emph{Social {Justice}
and {Political Change}: {Public Opinion} in {Capitalist} and
{Post-Communist States}}; Kluegel, J.R., Mason, D.S., Wegener, B., Eds.;
Walter de Gruyter: New York, 1995; pp. 239--259 ISBN 978-3-11-086894-4.}

\bibitem[\citeproctext]{ref-homan_being_2017}
\CSLLeftMargin{26. }%
\CSLRightInline{Homan, P.; Valentino, L.; Weed, E. Being and {Becoming
Poor}: {How Cultural Schemas Shape Beliefs About Poverty}. \emph{Social
Forces} \textbf{2017}, \emph{95}, 1023--1048,
doi:\href{https://doi.org/10.1093/sf/sox007}{10.1093/sf/sox007}.}

\bibitem[\citeproctext]{ref-castillo_legitimacy_2011}
\CSLLeftMargin{27. }%
\CSLRightInline{Castillo, J.C. Legitimacy of {Inequality} in a {Highly
Unequal Context}: {Evidence} from the {Chilean Case}. \emph{Social
Justice Research} \textbf{2011}, \emph{24}, 314--340,
doi:\href{https://doi.org/10.1007/s11211-011-0144-5}{10.1007/s11211-011-0144-5}.}

\bibitem[\citeproctext]{ref-evans_family_2010}
\CSLLeftMargin{28. }%
\CSLRightInline{Evans, M.; Kelley, J.; Sikora, J.; Treiman, D. Family
Scholarly Culture and Educational Success: {Books} and Schooling in 27
Nations. \emph{Research in Social Stratification and Mobility}
\textbf{2010}, \emph{28}, 171--197,
doi:\href{https://doi.org/10.1016/j.rssm.2010.01.002}{10.1016/j.rssm.2010.01.002}.}

\bibitem[\citeproctext]{ref-jasso_how_1999}
\CSLLeftMargin{29. }%
\CSLRightInline{Jasso, G. How {Much Injustice} Is {There} in the
{World}? {Two New Justice Indexes}. \emph{American Sociological Review}
\textbf{1999}, \emph{64}, 133--168,
doi:\href{https://doi.org/10.1177/000312249906400110}{10.1177/000312249906400110}.}

\bibitem[\citeproctext]{ref-shariff_income_2016}
\CSLLeftMargin{30. }%
\CSLRightInline{Shariff, A.F.; Wiwad, D.; Aknin, L.B. Income {Mobility
Breeds Tolerance} for {Income Inequality}: {Cross-National} and
{Experimental Evidence}. \emph{Perspectives on Psychological Science}
\textbf{2016}, \emph{11}, 373--380,
doi:\href{https://doi.org/10.1177/1745691616635596}{10.1177/1745691616635596}.}

\bibitem[\citeproctext]{ref-centeno_arc_2012}
\CSLLeftMargin{31. }%
\CSLRightInline{Centeno, M.A.; Cohen, J.N. The {Arc} of {Neoliberalism}.
\emph{Annual Review of Sociology} \textbf{2012}, \emph{38}, 317--340,
doi:\href{https://doi.org/10.1146/annurev-soc-081309-150235}{10.1146/annurev-soc-081309-150235}.}

\bibitem[\citeproctext]{ref-harvey_breve_2015}
\CSLLeftMargin{32. }%
\CSLRightInline{Harvey, D. \emph{{Breve historia del neoliberalismo}};
Ediciones Akal: Madrid (Espa{ñ}a), 2015; ISBN 978-84-460-2517-7.}

\bibitem[\citeproctext]{ref-jensen_worlds_2008}
\CSLLeftMargin{33. }%
\CSLRightInline{Jensen, C. Worlds of Welfare Services and Transfers.
\emph{Journal of European Social Policy} \textbf{2008}, \emph{18},
151--162,
doi:\href{https://doi.org/10.1177/0958928707087591}{10.1177/0958928707087591}.}

\bibitem[\citeproctext]{ref-stoy_worlds_2014}
\CSLLeftMargin{34. }%
\CSLRightInline{Stoy, V. Worlds of {Welfare Services}: {From Discovery}
to {Exploration}. \emph{Social Policy \& Administration} \textbf{2014},
\emph{48}, 343--360,
doi:\href{https://doi.org/10.1111/spol.12006}{10.1111/spol.12006}.}

\bibitem[\citeproctext]{ref-salamon_marketization_1993}
\CSLLeftMargin{35. }%
\CSLRightInline{Salamon, L.M. The {Marketization} of {Welfare}:
{Changing Nonprofit} and {For-Profit Roles} in the {American Welfare
State}. \emph{Social Service Review} \textbf{1993}, \emph{67}, 16--39,
doi:\href{https://doi.org/10.1086/603963}{10.1086/603963}.}

\bibitem[\citeproctext]{ref-sivesind_changing_2017}
\CSLLeftMargin{36. }%
\CSLRightInline{Sivesind, K.H.
\href{https://doi.org/10.1007/978-3-319-55381-8_2}{The {Changing Roles}
of {For-Profit} and {Nonprofit Welfare Provision} in {Norway}, {Sweden},
and {Denmark}}. In \emph{Promoting {Active Citizenship}}; Sivesind,
K.H., Saglie, J., Eds.; Springer International Publishing: Cham, 2017;
pp. 33--74 ISBN 978-3-319-55380-1 978-3-319-55381-8.}

\bibitem[\citeproctext]{ref-greitemeyer_subjective_2016}
\CSLLeftMargin{37. }%
\CSLRightInline{Greitemeyer, T.; Sagioglou, C. Subjective Socioeconomic
Status Causes Aggression: {A} Test of the Theory of Social Deprivation.
\emph{Journal of Personality and Social Psychology} \textbf{2016},
\emph{111}, 178--194,
doi:\href{https://doi.org/10.1037/pspi0000058}{10.1037/pspi0000058}.}

\bibitem[\citeproctext]{ref-mishra_subjective_2015}
\CSLLeftMargin{38. }%
\CSLRightInline{Mishra, S.; Carleton, R.N. Subjective Relative
Deprivation Is Associated with Poorer Physical and Mental Health.
\emph{Social Science \& Medicine} \textbf{2015}, \emph{147}, 144--149,
doi:\href{https://doi.org/10.1016/j.socscimed.2015.10.030}{10.1016/j.socscimed.2015.10.030}.}

\bibitem[\citeproctext]{ref-smith_relative_2012}
\CSLLeftMargin{39. }%
\CSLRightInline{Smith, H.J.; Pettigrew, T.F.; Pippin, G.M.;
Bialosiewicz, S. Relative {Deprivation}: {A Theoretical} and
{Meta-Analytic Review}. \emph{Personality and Social Psychology Review}
\textbf{2012}, \emph{16}, 203--232,
doi:\href{https://doi.org/10.1177/1088868311430825}{10.1177/1088868311430825}.}

\bibitem[\citeproctext]{ref-power_deprivationprotest_2018}
\CSLLeftMargin{40. }%
\CSLRightInline{Power, S.A. The {Deprivation-Protest Paradox}: {How} the
{Perception} of {Unfair Economic Inequality Leads} to {Civic Unrest}.
\emph{Current Anthropology} \textbf{2018}, \emph{59}, 765--789,
doi:\href{https://doi.org/10.1086/700679}{10.1086/700679}.}

\bibitem[\citeproctext]{ref-wilson_role_2003}
\CSLLeftMargin{41. }%
\CSLRightInline{Wilson, C. The {Role} of a {Merit Principle} in
{Distributive Justice}. \emph{The Journal of Ethics} \textbf{2003},
\emph{7}, 277--314,
doi:\href{https://doi.org/10.1023/A:1024667228488}{10.1023/A:1024667228488}.}

\bibitem[\citeproctext]{ref-bucca_merit_2016}
\CSLLeftMargin{42. }%
\CSLRightInline{Bucca, M. Merit and Blame in Unequal Societies:
{Explaining Latin Americans}' Beliefs about Wealth and Poverty.
\emph{Research in Social Stratification and Mobility} \textbf{2016},
\emph{44}, 98--112,
doi:\href{https://doi.org/10.1016/j.rssm.2016.02.005}{10.1016/j.rssm.2016.02.005}.}

\bibitem[\citeproctext]{ref-chong_mystery_2008}
\CSLLeftMargin{43. }%
\CSLRightInline{Chong, A.; Ñopo, H. The {Mystery} of {Discrimination} in
{Latin America}. \emph{Econom{í}a} \textbf{2008}, \emph{8}, 79--107,
doi:\href{https://doi.org/10.1353/eco.0.0005}{10.1353/eco.0.0005}.}

\bibitem[\citeproctext]{ref-torche_intergenerational_2014}
\CSLLeftMargin{44. }%
\CSLRightInline{Torche, F. Intergenerational {Mobility} and
{Inequality}: {The Latin American Case}. \emph{Annual Review of
Sociology} \textbf{2014}, \emph{40}, 619--642,
doi:\href{https://doi.org/10.1146/annurev-soc-071811-145521}{10.1146/annurev-soc-071811-145521}.}

\bibitem[\citeproctext]{ref-salgado_inequality_2023}
\CSLLeftMargin{45. }%
\CSLRightInline{Salgado, M.; Castillo, J. Inequality and
{Stratification} in {Latin America}. In \emph{The {Oxford Handbook} of
{Social Stratification}}; Gangl, M., Platt, L., Polavieja, J.G., van de
Werfhorst, H.G., Eds.; Oxford University Press: 2023, 2023; p. 0 ISBN
978-0-19-753948-4.}

\bibitem[\citeproctext]{ref-molyneux_neoliberal_2008}
\CSLLeftMargin{46. }%
\CSLRightInline{Molyneux, M. The {``{Neoliberal Turn}''} and the {New
Social Policy} in {Latin America}: {How Neoliberal}, {How New}?
\emph{Development and Change} \textbf{2008}, \emph{39}, 775--797,
doi:\href{https://doi.org/10.1111/j.1467-7660.2008.00505.x}{10.1111/j.1467-7660.2008.00505.x}.}

\bibitem[\citeproctext]{ref-centrodeestudiospublicos_estudio_2024}
\CSLLeftMargin{47. }%
\CSLRightInline{Centro de Estudios Públicos \emph{Estudio {Nacional} de
{Opini{ó}n P{ú}blica Base Consolidada}}; Centro de Estudios P{ú}blicos,
2024;}

\bibitem[\citeproctext]{ref-goldthorpe_myth_2003}
\CSLLeftMargin{48. }%
\CSLRightInline{Goldthorpe, J. The Myth of Education-Based Meritocracy.
\emph{New Economy} \textbf{2003}, \emph{10}, 234--239,
doi:\href{https://doi.org/10.1046/j.1468-0041.2003.00324.x}{10.1046/j.1468-0041.2003.00324.x}.}
